---
title: "2.5. Exploratory data analysis"
output:
  html_document:
    df_print: paged
---

```{r include = FALSE}
knitr::opts_chunk$set( warning = FALSE, message = FALSE)
library(janitor)
library(tidyverse)
library(lubridate)
library(ggplot2)

```

```{r}
raw_beds <- read_csv("data/beds_by_nhs_board_of_treatment_and_specialty.csv")

raw_activity <- read_csv("data/inpatient_and_daycase_by_nhs_board_of_treatment_and_specialty.csv")
```


# Cleaning Data
```{r}
wranged_beds_no_gj <- raw_beds %>% 
  clean_names() %>% 
  mutate(q = str_sub(quarter, -2)) %>% 
  # Change to date format
  mutate(quarter = yq(quarter)) %>% 
  # rename the columns above 
  mutate(date = quarter,
         quarter = q) %>% 
  # Change Heath Board names to read better
  mutate(hb = if_else(hb %in% "S08000015", "Ayrshire & Arran",
                      if_else(hb %in% "S08000016", "Borders",
                      if_else(hb %in% "S08000017", "Dumfries & Galloway",
                      if_else(hb %in% "S08000019", "Forth Valley",
                      if_else(hb %in% "S08000020", "Grampian",
                      if_else(hb %in% "S08000022", "Highland",
                      if_else(hb %in% "S08000024", "Lothian",
                      if_else(hb %in% "S08000025", "Orkney",
                      if_else(hb %in% "S08000026", "Shetland",
                      if_else(hb %in% "S08000028", "Western Isles",
                      if_else(hb %in% "S08000029", "Fife",
                      if_else(hb %in% "S08000030", "Tayside",
                      if_else(hb %in% "S08000031", "Greater Glasgow & Clyde",
                      if_else(hb %in% "S08000032", "Lanarkshire",
                      if_else(hb %in% "S92000003", "Scotland", NA_character_)
                      ))))))))))))))) %>% 
  # take out location names to take out duplicate
  filter(location_qf == "d") %>%
  
  # Take out of cleaning script and add when pulling in shiny
  
  # take out Golden Jubilee
  filter(hb != is.na(hb)) %>%
  # take out unused columns
  # take out location to remove duplicate
  select(date, quarter, hb, specialty_name, all_staffed_beds, 
         total_occupied_beds, average_available_staffed_beds,
         average_occupied_beds, percentage_occupancy)
  
wranged_beds_no_gj
```

```{r}
activity_no_gj <- raw_activity %>% 
  clean_names() %>% 
  # Create new column for quarters
  mutate(q = str_sub(quarter, -2)) %>% 
  # Change to date format
  mutate(quarter = yq(quarter)) %>% 
  # rename the columns above 
  mutate(date = quarter,
         quarter = q) %>% 
  # Change Heath Board names to read better
  mutate(
    hb = if_else(hb %in% "S08000015", "Ayrshire & Arran",
                  if_else(hb %in% "S08000016", "Borders",
                  if_else(hb %in% "S08000017", "Dumfries & Galloway",
                  if_else(hb %in% "S08000019", "Forth Valley",
                  if_else(hb %in% "S08000020", "Grampian",
                  if_else(hb %in% "S08000022", "Highland",
                  if_else(hb %in% "S08000024", "Lothian",
                  if_else(hb %in% "S08000025", "Orkney",
                  if_else(hb %in% "S08000026", "Shetland",
                  if_else(hb %in% "S08000028", "Western Isles",
                  if_else(hb %in% "S08000029", "Fife",
                  if_else(hb %in% "S08000030", "Tayside",
                  if_else(hb %in% "S08000031", "Greater Glasgow & Clyde",
                  if_else(hb %in% "S08000032", "Lanarkshire",
                  if_else(hb %in% "S92000003", "Scotland", NA_character_)
                  ))))))))))))))) %>% 
  # take out location names to take out duplicate
  filter(location_qf == "d") %>% 
  
  # Take out of cleaning script and add when pulling in shiny
  
  
  # take out Golden Jubilee
  filter(hb != is.na(hb)) %>%
  # take out unused columns
  select(date, quarter, hb, admission_type, 
         specialty, specialty_name, episodes, 
         length_of_episode, average_length_of_episode, 
         spells, length_of_spell, average_length_of_spell)
  
activity_no_gj 
```

# Exploratory analysis

```{r}
unique(activity_no_gj$specialty_name)
```


```{r}
unique(wranged_beds_no_gj$hb)
```
<br>

### KPI: Capacity – what is happening to the number of beds over the period? Perhaps think about the specialities these are if there has been specific variation?


Shows the percentage occupancy per Specialty grouping

```{r}
wranged_beds_no_gj %>% 
  filter(str_detect(specialty_name, " Grouping"),
         specialty_name != "Other Grouping",
         hb == "Scotland") %>% 
  mutate(specialty_name = str_sub(specialty_name, start = 1, end = -9)) %>%
  ggplot(aes(x = date, y = percentage_occupancy, col = specialty_name)) +
  geom_point() +
  geom_line()
```
<br>
Obstetrics and Dental look odd as they have 100% occupancy some of the time, some show seasonal peaks.

Can change the `head()` in the chunk below to filter down the specialty. For shiny can filter by the highest occupancy or the specialty_name
Have removed Obstetrics and Dental just to make clearer

```{r}
bigest_occupancy <- wranged_beds_no_gj %>% 
  filter(str_detect(specialty_name, " Grouping"),
         specialty_name != "Other Grouping",
         specialty_name != "Obstetrics Grouping",
         specialty_name != "Dental Grouping",
         hb == "Scotland") %>% 
  group_by(specialty_name) %>% 
  summarise(count = sum(percentage_occupancy)) %>% 
  arrange(desc(count)) %>% 
  # change head to 
  head(5)

wranged_beds_no_gj %>% 
  filter(specialty_name %in% c(bigest_occupancy$specialty_name)) %>%
  filter(str_detect(specialty_name, " Grouping"),
         specialty_name != "Other Grouping",
         hb == "Scotland") %>% 
  mutate(specialty_name = str_sub(specialty_name, start = 1, end = -9)) %>%
  ggplot(aes(x = date, y = percentage_occupancy, col = specialty_name)) +
  geom_point() +
  geom_line()
```
<br>
This shows the regions 

Can use for shiny changing the specialty_name.

Can add try to add a label to show what the highest number is

shows remote areas don't need/uses some services so have belive have less can over lay the number total_occupied_beds at high point to give scale


```{r}
wranged_beds_no_gj %>% 
  filter(str_detect(specialty_name, " Grouping"),
         specialty_name != "Other Grouping",
         specialty_name == "Emergency Grouping") %>% 
  mutate(specialty_name = str_sub(specialty_name, start = 1, end = -9)) %>%
  ggplot(aes(x = date, y = percentage_occupancy, col = hb)) +
  geom_point() +
  geom_line() +
  facet_wrap(~ hb) +
  theme(legend.position="none")
```
<br>

## seasonal trends for staffed beds in the top 3 specialty

Have taken the top 3 to show for ease. For shiny can filter by the highest occupancy or the specialty_name

```{r}
bigest_specialty <- wranged_beds_no_gj %>% 
  filter(str_detect(specialty_name, " Grouping"),
         specialty_name != "Other Grouping",
         hb == "Scotland") %>% 
  group_by(specialty_name) %>% 
  summarise(count = sum(all_staffed_beds)) %>% 
  arrange(desc(count)) %>% 
  head(3)


wranged_beds_no_gj %>% 
  filter(specialty_name %in% c(bigest_specialty$specialty_name)) %>% 
  filter(str_detect(specialty_name, " Grouping"),
         specialty_name != "Other Grouping",
         hb == "Scotland") %>% 
  mutate(specialty_name = str_sub(specialty_name, start = 1, end = -9)) %>%
  ggplot(aes(x = date, y = all_staffed_beds, col = specialty_name)) +
  geom_point() +
  geom_line()
```
<br>
These two are more general shorts of the bed occupancy over Scotland. Can do a tab for both of them togeter

Shows the seasonal peaks in occupancy in Q1 & Q4
Also show a steady decrease in the number of staffed beds from 2016 to 2019

```{r}
wranged_beds_no_gj %>% 
  filter(specialty_name %in% c("All Acute", "All Specialties"),
         hb %in% "Scotland") %>% 
  ggplot(aes(x = date, y = percentage_occupancy, col = specialty_name)) +
  geom_point() +
  geom_line()
```

```{r}
wranged_beds_no_gj %>% 
  filter(specialty_name %in% c("All Acute", "All Specialties"),
         hb %in% "Scotland") %>%  
  ggplot(aes(x = date, y = average_occupied_beds, col = specialty_name)) +
  # geom_col(position = "dodge")
  geom_point() +
  geom_line()
```
<br>
Shows the regions brake down 

Glasgow is the hub for most types of beds

```{r}
wranged_beds_no_gj %>% 
  filter(specialty_name %in% c("All Acute", "All Specialties"),
         hb != "Scotland") %>%
  ggplot(aes(x = date, y = all_staffed_beds, col = specialty_name)) +
  geom_point() +
  geom_line() +
  facet_wrap(~ hb) +
  theme(legend.position="none")
```

```{r}
wranged_beds_no_gj %>% 
  filter(specialty_name %in% c("All Acute", "All Specialties"),
         hb != "Scotland") %>%
  ggplot(aes(x = date, y = percentage_occupancy, col = specialty_name)) +
  geom_point() +
  geom_line() +
  facet_wrap(~ hb) +
  theme(legend.position="none")
```



<br>

Can use this in shiny just filter it down by specialty

Can have both elective or emergency admissions show between the two graphs 

```{r}
activity_no_gj %>%  
  filter(hb %in% "Scotland",
         admission_type %in% "Emergency Inpatients") %>% 
  ggplot(aes(x = date, y = episodes, col = specialty_name)) +
  geom_point() +
  geom_line()
```

```{r}
activity_no_gj %>%  
  filter(hb %in% "Scotland",
         admission_type %in% "Emergency Inpatients",
         specialty_name %in% "General Medicine") %>% 
 # group_by(episodes) %>% 
  ggplot(aes(x = date, y = episodes, col = specialty_name)) +
  geom_point() +
  geom_line()
```


```{r}
activity_no_gj %>%  
  filter(hb %in% "Scotland",
         admission_type %in% "Elective Inpatients") %>%
  ggplot(aes(x = date, y = episodes, col = specialty_name)) +
  geom_point() +
  geom_line()
```
<br>

### Admission by specialty

Better can shiny this one and pick the specialty_name

#### KPI: Do people have longer lengths of stay from elective or emergency admissions?

### KPI: The whole system: from what’s coming through the front door in A&E and then how it filters through; how many emergency admissions are there, how many electives, how does this impact the lengths of stay people have in hospital?

take out "All Inpatients and Day cases" & "Not Specified" as duplicating and not really any there

##### Episodes by Admission
```{r}
activity_no_gj %>%  
  filter(hb %in% "Scotland",
         specialty_name %in% "Infectious Diseases",
         admission_type %in% c("Elective Inpatients", 
                               "Emergency Inpatients", 
                               "Transfers")) %>% 
  ggplot(aes(x = date, y = episodes, col = admission_type)) +
  geom_point() +
  geom_line()
```
<br>
##### Spells by Admission
```{r}
activity_no_gj %>%  
  filter(hb %in% "Scotland",
         specialty_name %in% "Infectious Diseases",
         admission_type %in% c("Elective Inpatients", 
                               "Emergency Inpatients", 
                               "Transfers")) %>% 
  ggplot(aes(x = date, y = spells, col = admission_type)) +
  geom_point() +
  geom_line()
```
<br>
Can show by regions brake down 

Tayside had a larger bump in Emergency Inpatients durung the pandemic  "Infectious Diseases"



```{r}
activity_no_gj %>%  
  filter(hb != "Scotland",
         specialty_name %in% "Infectious Diseases",
         admission_type %in% c("Elective Inpatients", 
                               "Emergency Inpatients", 
                               "Transfers")) %>% 
  ggplot(aes(x = date, y = spells, col = admission_type)) +
  geom_point() +
  geom_line() +
  facet_wrap(~ hb) +
  theme(legend.position="none")
```




```{r}
activity_no_gj %>%  
  filter(hb != "Scotland",
         specialty_name %in% "General Surgery",
         admission_type %in% c("Elective Inpatients", 
                               "Emergency Inpatients", 
                               "Transfers")) %>% 
  ggplot(aes(x = date, y = episodes, col = admission_type)) +
  geom_point() +
  geom_line() +
  facet_wrap(~ hb) +
  theme(legend.position="none")
```
<br>
"Endocrinology & Diabetes"
"Emergency Inpatients"
   fife has a large ingress then decrees
   Lothian is very seasonal summer 
   
"Medical Oncology
  Much be three areas of treatment
  Greater Glasgow & Clyde similar Elective Inpatients, Emergency Inpatients 
  Grampian 
  Lothian high amount

```{r}
unique(activity_no_gj$admission_type)
```


```{r}
unique(activity_no_gj$specialty_name)
```

```{r}
unique(activity_no_gj$hb)
```
<br>
<br>
<br>