---
title: "R Notebook"
output: html_notebook
---


```{r}
library(janitor)
library(tidyverse)
library(lubridate)
library(ggplot2)

raw_beds <- read_csv("data/beds_by_nhs_board_of_treatment_and_specialty.csv")
```


```{r}
wranged_beds_no_gj <- raw_beds %>% 
  clean_names() %>% 
  mutate(q = str_sub(quarter, -2)) %>% 
  # Change to date format
  mutate(quarter = yq(quarter)) %>% 
  # rename the columns above 
  mutate(date = quarter,
         quarter = q) %>% 
  # Change Heath Board names to read better
  mutate(hb = if_else(hb %in% "S08000015", "Ayrshire & Arran",
                      if_else(hb %in% "S08000016", "Borders",
                      if_else(hb %in% "S08000017", "Dumfries & Galloway",
                      if_else(hb %in% "S08000019", "Forth Valley",
                      if_else(hb %in% "S08000020", "Grampian",
                      if_else(hb %in% "S08000022", "Highland",
                      if_else(hb %in% "S08000024", "Lothian",
                      if_else(hb %in% "S08000025", "Orkney",
                      if_else(hb %in% "S08000026", "Shetland",
                      if_else(hb %in% "S08000028", "Western Isles",
                      if_else(hb %in% "S08000029", "Fife",
                      if_else(hb %in% "S08000030", "Tayside",
                      if_else(hb %in% "S08000031", "Greater Glasgow & Clyde",
                      if_else(hb %in% "S08000032", "Lanarkshire",
                      if_else(hb %in% "S92000003", "Scotland", NA_character_)
                      ))))))))))))))) %>% 
  # take out location names to take out duplicate
  filter(location_qf == "d") %>%
  # take out Golden Jubilee
  filter(hb != is.na(hb)) %>%
  # take out unused columns
  # take out location to remove duplicate
  select(date, quarter, hb, specialty_name, all_staffed_beds, 
         total_occupied_beds, average_available_staffed_beds,
         average_occupied_beds, percentage_occupancy)
  
wranged_beds_no_gj
```

### KPI: Capacity – what is happening to the number of beds over the period? Perhaps think about the specialities these are if there has been specific variation?


Shows the percentage occupancy per Specialty grouping

```{r}
wranged_beds_no_gj %>% 
  filter(str_detect(specialty_name, " Grouping"),
         specialty_name != "Other Grouping",
         hb == "Scotland") %>% 
  mutate(specialty_name = str_sub(specialty_name, start = 1, end = -9)) %>%
  ggplot(aes(x = date, y = percentage_occupancy, col = specialty_name)) +
  geom_point() +
  geom_line()
```
Obstetrics and Dental look odd as they have 100% occupancy some of the time

```{r}
bigest_occupancy <- wranged_beds_no_gj %>% 
  filter(str_detect(specialty_name, " Grouping"),
         specialty_name != "Other Grouping",
         specialty_name != "Obstetrics Grouping",
         specialty_name != "Dental Grouping",
         hb == "Scotland") %>% 
  group_by(specialty_name) %>% 
  summarise(count = sum(percentage_occupancy)) %>% 
  arrange(desc(count)) %>% 
  head(3)

wranged_beds_no_gj %>% 
  filter(specialty_name %in% c(bigest_occupancy$specialty_name)) %>%
  filter(str_detect(specialty_name, " Grouping"),
         specialty_name != "Other Grouping",
         hb == "Scotland") %>% 
  mutate(specialty_name = str_sub(specialty_name, start = 1, end = -9)) %>%
  ggplot(aes(x = date, y = percentage_occupancy, col = specialty_name)) +
  geom_point() +
  geom_line()
```


Can use for shiny changing the specialty_name.

shows remote areas don't need/uses some services so have belive have less can over lay the number total_occupied_beds at high point to give scale


```{r}
wranged_beds_no_gj %>% 
  filter(str_detect(specialty_name, " Grouping"),
         specialty_name != "Other Grouping",
         specialty_name == "Emergency Grouping") %>% 
  mutate(specialty_name = str_sub(specialty_name, start = 1, end = -9)) %>%
  ggplot(aes(x = date, y = percentage_occupancy, col = hb)) +
  geom_point() +
  geom_line() +
  facet_wrap(~ hb) +
  theme(legend.position="none")
```


## seasonal trends for staffed beds in the top 3 specialty

Have taken the top 3 to show for ease

```{r}
bigest_specialty <- wranged_beds_no_gj %>% 
  filter(str_detect(specialty_name, " Grouping"),
         specialty_name != "Other Grouping",
         hb == "Scotland") %>% 
  group_by(specialty_name) %>% 
  summarise(count = sum(all_staffed_beds)) %>% 
  arrange(desc(count)) %>% 
  head(3)


wranged_beds_no_gj %>% 
  filter(specialty_name %in% c(bigest_specialty$specialty_name)) %>% 
  filter(str_detect(specialty_name, " Grouping"),
         specialty_name != "Other Grouping",
         hb == "Scotland") %>% 
  mutate(specialty_name = str_sub(specialty_name, start = 1, end = -9)) %>%
  ggplot(aes(x = date, y = all_staffed_beds, col = specialty_name)) +
  geom_point() +
  geom_line()
```



```{r}
wranged_beds_no_gj %>% 
  filter(specialty_name %in% c("All Acute", "All Specialties"),
         hb %in% "Scotland") %>% 
  ggplot(aes(x = date, y = percentage_occupancy, col = specialty_name)) +
  geom_point() +
  geom_line()
```

```{r}
wranged_beds_no_gj %>% 
  filter(specialty_name %in% c("All Acute", "All Specialties"),
         hb %in% "Scotland") %>% 
  ggplot(aes(x = date, y = all_staffed_beds, col = specialty_name)) +
  geom_point() +
  geom_line()
```

Shows Glasgow is the hub for most 

```{r}
wranged_beds_no_gj %>% 
  filter(specialty_name %in% c("All Acute", "All Specialties"),
         hb != "Scotland") %>%
  ggplot(aes(x = date, y = all_staffed_beds, col = specialty_name)) +
  geom_point() +
  geom_line() +
  facet_wrap(~ hb) +
  theme(legend.position="none")
```