---
title: "R Notebook"
output: html_notebook
---

```{r}
library(feasts)
```


```{r}
library(janitor)
library(tidyverse)
library(lubridate)
library(ggplot2)

raw_beds <- read_csv("data/beds_by_nhs_board_of_treatment_and_specialty.csv")
```

```{r}
glimpse(raw_beds)
```

```{r}
wranged_beds_no_gj <- raw_beds %>% 
  clean_names() %>% 
  mutate(q = str_sub(quarter, -2)) %>% 
  # Change to date format
  mutate(quarter = yq(quarter)) %>% 
  # rename the columns above 
  mutate(date = quarter,
         quarter = q) %>% 
  # Change Heath Board names to read better
  mutate(hb = if_else(hb %in% "S08000015", "Ayrshire & Arran",
                      if_else(hb %in% "S08000016", "Borders",
                      if_else(hb %in% "S08000017", "Dumfries & Galloway",
                      if_else(hb %in% "S08000019", "Forth Valley",
                      if_else(hb %in% "S08000020", "Grampian",
                      if_else(hb %in% "S08000022", "Highland",
                      if_else(hb %in% "S08000024", "Lothian",
                      if_else(hb %in% "S08000025", "Orkney",
                      if_else(hb %in% "S08000026", "Shetland",
                      if_else(hb %in% "S08000028", "Western Isles",
                      if_else(hb %in% "S08000029", "Fife",
                      if_else(hb %in% "S08000030", "Tayside",
                      if_else(hb %in% "S08000031", "Greater Glasgow & Clyde",
                      if_else(hb %in% "S08000032", "Lanarkshire",
                      if_else(hb %in% "S92000003", "Scotland", NA_character_)
                      ))))))))))))))) %>% 
  # take out location names to take out duplicate
  filter(location_qf == "d") %>%
  # take out Golden Jubilee
  filter(hb != is.na(hb)) %>%
  # take out unused columns
  # take out location to remove duplicate
  select(date, quarter, hb, specialty_name, all_staffed_beds, 
         total_occupied_beds, average_available_staffed_beds,
         average_occupied_beds, percentage_occupancy)
  
wranged_beds_no_gj
```


```{r}
wranged_beds_with_gj <- raw_beds %>% 
  clean_names() %>% 
  mutate(q = str_sub(quarter, -2)) %>% 
  # Change to date format
  mutate(quarter = yq(quarter)) %>% 
  # rename the columns above 
  mutate(date = quarter,
         quarter = q) %>% 
  # Change Heath Board names to read better
  mutate(hb = if_else(hb %in% "SB0801", "Golden Jubilee",
                      if_else(hb %in% "S08000015", "Ayrshire & Arran",
                      if_else(hb %in% "S08000016", "Borders",
                      if_else(hb %in% "S08000017", "Dumfries & Galloway",
                      if_else(hb %in% "S08000019", "Forth Valley",
                      if_else(hb %in% "S08000020", "Grampian",
                      if_else(hb %in% "S08000022", "Highland",
                      if_else(hb %in% "S08000024", "Lothian",
                      if_else(hb %in% "S08000025", "Orkney",
                      if_else(hb %in% "S08000026", "Shetland",
                      if_else(hb %in% "S08000028", "Western Isles",
                      if_else(hb %in% "S08000029", "Fife",
                      if_else(hb %in% "S08000030", "Tayside",
                      if_else(hb %in% "S08000031", "Greater Glasgow & Clyde",
                      if_else(hb %in% "S08000032", "Lanarkshire",
                      if_else(hb %in% "S92000003", "Scotland", hb)
                      )))))))))))))))) %>% 
  # take out location names to take out duplicate
  filter(location_qf == "d") %>% 
  # take out unused columns
  # take out location to remove duplicate
  select(date, quarter, hb, specialty_name, all_staffed_beds, 
         total_occupied_beds, average_available_staffed_beds,
         average_occupied_beds, percentage_occupancy)
  
wranged_beds_with_gj
```


```{r}
# wranged_beds <- raw_beds %>% 
#   clean_names() %>% 
#   mutate(q = str_sub(quarter, -2)) %>% 
#   # Change to date format
#   mutate(quarter = yq(quarter)) %>% 
#   # rename the columns above 
#   mutate(date = quarter,
#          quarter = q) %>% 
#   # Change Heath Board names to read better
#     # probly don'tneed
#   mutate(hb = if_else(hb %in% "SB0801", "Golden Jubilee",
#                       if_else(hb %in% "S08000015", "Ayrshire & Arran",
#                       if_else(hb %in% "S08000016", "Borders",
#                       if_else(hb %in% "S08000017", "Dumfries & Galloway",
#                       if_else(hb %in% "S08000019", "Forth Valley",
#                       if_else(hb %in% "S08000020", "Grampian",
#                       if_else(hb %in% "S08000022", "Highland",
#                       if_else(hb %in% "S08000024", "Lothian",
#                       if_else(hb %in% "S08000025", "Orkney",
#                       if_else(hb %in% "S08000026", "Shetland",
#                       if_else(hb %in% "S08000028", "Western Isles",
#                       if_else(hb %in% "S08000029", "Fife",
#                       if_else(hb %in% "S08000030", "Tayside",
#                       if_else(hb %in% "S08000031", "Greater Glasgow & Clyde",
#                       if_else(hb %in% "S08000032", "Lanarkshire",
#                       if_else(hb %in% "S92000003", "Scotland", hb)
#                       )))))))))))))))) %>% 
#   # Change location names to take out dupitice
#   filter(location_qf == "d") %>% 
#   # mutate(
#   #   location = if_else(location %in% "SB0801", "Golden Jubilee",
#   #                 if_else(location %in% "S08000015", "Ayrshire & Arran",
#   #                 if_else(location %in% "S08000016", "Borders",
#   #                 if_else(location %in% "S08000017", "Dumfries & Galloway",
#   #                 if_else(location %in% "S08000019", "Forth Valley",
#   #                 if_else(location %in% "S08000020", "Grampian",
#   #                 if_else(location %in% "S08000022", "Highland",
#   #                 if_else(location %in% "S08000024", "Lothian",
#   #                 if_else(location %in% "S08000025", "Orkney",
#   #                 if_else(location %in% "S08000026", "Shetland",
#   #                 if_else(location %in% "S08000028", "Western Isles",
#   #                 if_else(location %in% "S08000029", "Fife",
#   #                 if_else(location %in% "S08000030", "Tayside",
#   #                 if_else(location %in% "S08000031", "Greater Glasgow & Clyde",
#   #                 if_else(location %in% "S08000032", "Lanarkshire",
#   #                 if_else(location %in% "S92000003", "Scotland", NA_character_ 
#   #                 ))))))))))))))))) %>% 
#   # filter(location != is.na(location)) %>% 
#   # take out unused columns
#   # take out location to remove duplicate
#   select(date, quarter, hb, specialty_name, all_staffed_beds, 
#          total_occupied_beds, average_available_staffed_beds,
#          average_occupied_beds, percentage_occupancy)
#   
# wranged_beds 
```

```{r}
# beds <- wranged_beds %>% 
#   mutate(
#     specialty_name = if_else(specialty_name %in% "+ Emergency +","A & E",
#                       if_else(specialty_name %in% "Paediatric", "Paediatric",
#                       if_else(specialty_name %in% "Surgery", "Surgery",
#                       if_else(specialty_name %in% "Newborn", "Midwifery", 
#                               specialty_name)))))%>% 
#   count(specialty_name)

```


```{r}
wranged_beds_no_gj %>% 
  count(hb)

wranged_beds_no_gj %>% 
  filter(str_detect(specialty_name, " Grouping")) %>% 
  count(specialty_name)

# wranged_beds_with_gj %>% 
#   count(location)
```

```{r}
wranged_beds_no_gj %>% 
  filter(str_detect(specialty_name, " Grouping"),
         specialty_name != "Other Grouping",
         hb != "Scotland") %>% 
  mutate(specialty_name = str_sub(specialty_name, start = 1, end = -9)) %>%
  ggplot(aes(x = date, y = percentage_occupancy, col = specialty_name)) +
  geom_point() +
  geom_line()+
  facet_wrap(~ hb)
```

```{r}
wranged_beds_no_gj %>% 
  filter(str_detect(specialty_name, " Grouping"),
         specialty_name != "Other Grouping",
         hb == "Scotland") %>% 
  mutate(specialty_name = str_sub(specialty_name, start = 1, end = -9)) %>%
  ggplot(aes(x = date, y = percentage_occupancy, col = specialty_name)) +
  geom_point() +
  geom_line()
```

```{r}
wranged_beds_no_gj %>% 
  filter(hb == "Scotland",
         specialty_name == "Obstetrics Grouping")
```

```{r}
wranged_beds_no_gj %>% 
  filter(hb == "Scotland",
         specialty_name == "Dental Grouping")
```


```{r}
unique(wranged_beds_no_gj$specialty_name)
```

```{r}
wranged_beds_no_gj_q <- raw_beds %>% 
  clean_names() %>% 
  # (q = str_sub(quarter, -2)) %>% 
  # Change to date format
  mutate(quarter = yq(quarter)) %>% 
  # rename the columns above 
  mutate(date = quarter
         # ,
         # quarter = q
         ) %>%
  # Change Heath Board names to read better
  mutate(hb = if_else(hb %in% "S08000015", "Ayrshire & Arran",
                      if_else(hb %in% "S08000016", "Borders",
                      if_else(hb %in% "S08000017", "Dumfries & Galloway",
                      if_else(hb %in% "S08000019", "Forth Valley",
                      if_else(hb %in% "S08000020", "Grampian",
                      if_else(hb %in% "S08000022", "Highland",
                      if_else(hb %in% "S08000024", "Lothian",
                      if_else(hb %in% "S08000025", "Orkney",
                      if_else(hb %in% "S08000026", "Shetland",
                      if_else(hb %in% "S08000028", "Western Isles",
                      if_else(hb %in% "S08000029", "Fife",
                      if_else(hb %in% "S08000030", "Tayside",
                      if_else(hb %in% "S08000031", "Greater Glasgow & Clyde",
                      if_else(hb %in% "S08000032", "Lanarkshire",
                      if_else(hb %in% "S92000003", "Scotland", NA_character_)
                      ))))))))))))))) %>% 
  # take out location names to take out duplicate
  filter(location_qf == "d") %>%
  # take out Golden Jubilee
  filter(hb != is.na(hb)) %>%
  # take out unused columns
  # take out location to remove duplicate
  select(date,  hb, specialty_name, all_staffed_beds, 
         total_occupied_beds, average_available_staffed_beds,
         average_occupied_beds, percentage_occupancy)
  
wranged_beds_no_gj_q
```
quarter,

```{r}
wranged_beds_no_gj_q %>% 
  filter(str_detect(specialty_name, " Grouping"),
         specialty_name != "Other Grouping",
         hb %in% "Scotland",
         specialty_name == "Emergency Grouping") %>%
  mutate(specialty_name = str_sub(specialty_name, start = 1, end = -9)) %>%
  as_tsibble(index = date) %>%
  gg_season(percentage_occupancy)
```


```{r}
wranged_beds_no_gj %>% 
  filter(specialty_name %in% "All Acute") %>% 
  ggplot(aes(x = date, y = percentage_occupancy, col = hb)) +
  geom_point() +
  geom_line()+
  facet_wrap(~ hb) +
  theme(legend.position="none")
```

```{r}
wranged_beds_no_gj %>% 
  filter(specialty_name %in% "All Specialties") %>% 
  ggplot(aes(x = date, y = percentage_occupancy, col = hb)) +
  geom_point() +
  geom_line() +
  facet_wrap(~ hb) +
  theme(legend.position="none")
```

```{r}
wranged_beds_no_gj %>% 
  filter(specialty_name %in% c("All Acute", "All Specialties"),
         hb %in% "Scotland") %>% 
  ggplot(aes(x = date, y = percentage_occupancy, col = specialty_name)) +
  geom_point() +
  geom_line()
```

```{r}
wranged_beds_no_gj %>% 
  filter(specialty_name %in% c("All Acute", "All Specialties"),
         hb %in% "Scotland") %>% 
  ggplot(aes(x = date, y = all_staffed_beds, col = specialty_name)) +
  geom_point() +
  geom_line()
```


```{r}
raw_beds %>%
  clean_names() %>% 
  filter(specialty_name %in% "All Acute",
         hb %in% "S08000028")
```

```{r}
wranged_beds_no_gj %>% 
  filter(specialty_name %in% c("All Acute", "All Specialties"),
         hb != "Scotland",
         hb != "Golden Jubilee") %>%
  ggplot(aes(x = date, y = all_staffed_beds, col = specialty_name)) +
  geom_point() +
  geom_line() +
  facet_wrap(~ hb) +
  theme(legend.position="none")
```



```{r}
# line_col <- c(all_acute = "red", all_specialties = "green")
# 
# red <- wranged_beds %>% 
#   filter(specialty_name %in% "All Acute",
#          hb != "Scotland",
#          hb != "Golden Jubilee") %>% 
#   #group_by(specialty_name) %>% 
#   # pivot_wider(names_from = specialty_name, 
#   #             values_from = specialty_name) %>% 
#   # clean_names() %>% 
#   ggplot(aes(x = date, y = all_staffed_beds)) +
#   geom_point(aes(col = hb)) +
#   geom_line(aes(group = hb, col = "red")) +
#   facet_wrap(~ hb) +
#   theme(legend.position="none")
# 
# green <- wranged_beds %>% 
#   filter(specialty_name %in% "All Specialties",
#          hb != "Scotland",
#          hb != "Golden Jubilee") %>% 
#   #group_by(specialty_name) %>% 
#   # pivot_wider(names_from = specialty_name, 
#   #             values_from = specialty_name) %>% 
#   # clean_names() %>% 
#   ggplot(aes(x = date, y = all_staffed_beds)) +
#   geom_point(aes(col = hb)) +
#   geom_line(aes(group = hb, col = "green")) +
#   facet_wrap(~ hb) +
#   theme(legend.position="none")
# 
# cowplot::plot_grid(red, green) 
```



```{r}
wranged_beds_no_gj %>% 
  filter(specialty_name %in% c("All Acute"),
         hb %in% "Scotland") %>% 
  ggplot(aes(x = quarter, y = all_staffed_beds)) +
  geom_point() +
  geom_line(aes(y = ave(all_staffed_beds))) +
  #geom_smooth(aes(fill = specialty_name, color = specialty_name))
  geom_ribbon(aes(ymin = all_staffed_beds - 1, ymax = all_staffed_beds + 1), fill = "grey70", orientation = "x") 
  
```

use gg_sesoins
take out the 
find min/max and avg for each quarter



https://www.rspatial.org/intr/12-plots.html
Multiple plots

To combine multiple plots in one image you can use layout() or par(). For example, 4 plots can be combined with layout(matrix(1:4, 2, 2)) or par(mfrow=c(2,2)).

par(mfrow=c(2,2))
boxplot(iris$Sepal.Length~iris$Species, main = "Boxplot of Sepal Length",
        xlab = "Species", ylab = "Sepal Length", col = c("red","green3","blue"),
        cex.lab = 1.25)
hist(iris$Sepal.Length, main = "Histogram of Sepal Length",
     xlab = "Sepal Length", ylab = "Frequency", col = c("grey"), cex.lab = 1.25)
plot(iris$Sepal.Length, type = 'l', lty = 2, lwd = 2, col = 'red',
     main = "Variation with of Petal Width and Petal Length",
     xlab = "Petal Length", ylab = "Petal Width", cex.lab = 1.25)
plot(iris$Petal.Length, iris$Petal.Width, pch = 8, col = mycolor,
     main = "Scatter plot of Petal Width and Petal Length",
     xlab = "Petal Length", ylab = "Petal Width", cex.lab = 1.25)
