---
title: "R Notebook"
output: html_notebook
---

```{r}
library(janitor)
library(tidyverse)
library(lubridate)
library(ggplot2)
```


```{r}
#raw_staff_beds <- read_csv("data/beds_by_nhs_board_of_treatment_and_specialty.csv")
raw_activity <- read_csv("data/inpatient_and_daycase_by_nhs_board_of_treatment_and_specialty.csv")
```
```{r}
head(raw_activity)
```

```{r}
# activity <- raw_activity %>% 
#   clean_names() %>% 
#   # Create new column for quarters
#   mutate(q = str_sub(quarter, -2)) %>% 
#   # Change to date format
#   mutate(quarter = yq(quarter)) %>% 
#   # rename the columns above 
#   mutate(date = quarter,
#          quarter = q) %>% 
#   # Change Heath Board names to read better
#   mutate(
#     hb = if_else(hb %in% "SB0801", "Golden Jubilee",
#                   if_else(hb %in% "S08000015", "Ayrshire & Arran",
#                   if_else(hb %in% "S08000016", "Borders",
#                   if_else(hb %in% "S08000017", "Dumfries & Galloway",
#                   if_else(hb %in% "S08000019", "Forth Valley",
#                   if_else(hb %in% "S08000020", "Grampian",
#                   if_else(hb %in% "S08000022", "Highland",
#                   if_else(hb %in% "S08000024", "Lothian",
#                   if_else(hb %in% "S08000025", "Orkney",
#                   if_else(hb %in% "S08000026", "Shetland",
#                   if_else(hb %in% "S08000028", "Western Isles",
#                   if_else(hb %in% "S08000029", "Fife",
#                   if_else(hb %in% "S08000030", "Tayside",
#                   if_else(hb %in% "S08000031", "Greater Glasgow & Clyde",
#                   if_else(hb %in% "S08000032", "Lanarkshire",
#                   if_else(hb %in% "S92000003", "Scotland", hb)
#                   )))))))))))))))) %>% 
#   # take out location names to take out duplicate
#   filter(location_qf == "d") %>% 
#   # take out unsed columns
#   select(date, quarter, hb, admission_type, 
#          specialty, specialty_name, episodes, 
#          length_of_episode, average_length_of_episode, 
#          spells, length_of_spell, average_length_of_spell)
#   
# activity 
```


```{r}
activity_no_gj <- raw_activity %>% 
  clean_names() %>% 
  # Create new column for quarters
  mutate(q = str_sub(quarter, -2)) %>% 
  # Change to date format
  mutate(quarter = yq(quarter)) %>% 
  # rename the columns above 
  mutate(date = quarter,
         quarter = q) %>% 
  # Change Heath Board names to read better
  mutate(
    hb = if_else(hb %in% "S08000015", "Ayrshire & Arran",
                  if_else(hb %in% "S08000016", "Borders",
                  if_else(hb %in% "S08000017", "Dumfries & Galloway",
                  if_else(hb %in% "S08000019", "Forth Valley",
                  if_else(hb %in% "S08000020", "Grampian",
                  if_else(hb %in% "S08000022", "Highland",
                  if_else(hb %in% "S08000024", "Lothian",
                  if_else(hb %in% "S08000025", "Orkney",
                  if_else(hb %in% "S08000026", "Shetland",
                  if_else(hb %in% "S08000028", "Western Isles",
                  if_else(hb %in% "S08000029", "Fife",
                  if_else(hb %in% "S08000030", "Tayside",
                  if_else(hb %in% "S08000031", "Greater Glasgow & Clyde",
                  if_else(hb %in% "S08000032", "Lanarkshire",
                  if_else(hb %in% "S92000003", "Scotland", NA_character_)
                  ))))))))))))))) %>% 
  # take out location names to take out duplicate
  filter(location_qf == "d") %>% 
  # take out Golden Jubilee
  filter(hb != is.na(hb)) %>%
  # take out unsed columns
  select(date, quarter, hb, admission_type, 
         specialty, specialty_name, episodes, 
         length_of_episode, average_length_of_episode, 
         spells, length_of_spell, average_length_of_spell)
  
activity_no_gj 
```


```{r}
# activity2 <- raw_activity %>% 
#   clean_names() %>% 
#   # Create new column for quarters
#   mutate(q = str_sub(quarter, -2)) %>% 
#   # Change to date format
#   mutate(quarter = yq(quarter)) %>% 
#   # rename the columns above 
#   mutate(date = quarter,
#          quarter = q) %>% 
#   # Change Heath Board names to read better
#     # probly don't need
#   mutate(
#     hb = if_else(hb %in% "SB0801", "Golden Jubilee",
#                   if_else(hb %in% "S08000015", "Ayrshire & Arran",
#                   if_else(hb %in% "S08000016", "Borders",
#                   if_else(hb %in% "S08000017", "Dumfries & Galloway",
#                   if_else(hb %in% "S08000019", "Forth Valley",
#                   if_else(hb %in% "S08000020", "Grampian",
#                   if_else(hb %in% "S08000022", "Highland",
#                   if_else(hb %in% "S08000024", "Lothian",
#                   if_else(hb %in% "S08000025", "Orkney",
#                   if_else(hb %in% "S08000026", "Shetland",
#                   if_else(hb %in% "S08000028", "Western Isles",
#                   if_else(hb %in% "S08000029", "Fife",
#                   if_else(hb %in% "S08000030", "Tayside",
#                   if_else(hb %in% "S08000031", "Greater Glasgow & Clyde",
#                   if_else(hb %in% "S08000032", "Lanarkshire",
#                   if_else(hb %in% "S92000003", "Scotland", hb)
#                   )))))))))))))))) %>% 
#   # Change location names to take out dupitice
#   mutate(
#     location = if_else(location %in% "SB0801", "Golden Jubilee",
#                   if_else(location %in% "S08000015", "Ayrshire & Arran",
#                   if_else(location %in% "S08000016", "Borders",
#                   if_else(location %in% "S08000017", "Dumfries & Galloway",
#                   if_else(location %in% "S08000019", "Forth Valley",
#                   if_else(location %in% "S08000020", "Grampian",
#                   if_else(location %in% "S08000022", "Highland",
#                   if_else(location %in% "S08000024", "Lothian",
#                   if_else(location %in% "S08000025", "Orkney",
#                   if_else(location %in% "S08000026", "Shetland",
#                   if_else(location %in% "S08000028", "Western Isles",
#                   if_else(location %in% "S08000029", "Fife",
#                   if_else(location %in% "S08000030", "Tayside",
#                   if_else(location %in% "S08000031", "Greater Glasgow & Clyde",
#                   if_else(location %in% "S08000032", "Lanarkshire",
#                   if_else(location %in% "S92000003", "Scotland", NA_character_ 
#                         #  if_else(location %in% "^[A-Z0-9]{5}", , location)
#                         # couldn't get the Regex to work 
#                   ))))))))))))))))) %>% 
#   # filter out the na
#   filter(location != is.na(location)) %>% 
#   # take out unsed columns
#   select(date, quarter, hb, admission_type, 
#          specialty, specialty_name, episodes, 
#          length_of_episode, average_length_of_episode, 
#          spells, length_of_spell, average_length_of_spell)
#   
# activity2 
```

```{r}
library(compare)
```



```{r}
comparison <- compare(activity,activity2, equal = FALSE)
comparison$tM
```


Can use this in shiny just filter it down by specialty

Can have both elective or emergency admissions show between the two graphs 

```{r}
activity_no_gj %>%  
  filter(hb %in% "Scotland",
         admission_type %in% "Emergency Inpatients") %>% 
  ggplot(aes(x = date, y = episodes, col = specialty_name)) +
  geom_point() +
  geom_line()
```

```{r}
activity_no_gj %>%  
  filter(hb %in% "Scotland",
         admission_type %in% "Emergency Inpatients",
         specialty_name %in% "General Medicine") %>% 
 # group_by(episodes) %>% 
  ggplot(aes(x = date, y = episodes, col = specialty_name)) +
  geom_point() +
  geom_line()
```


```{r}
activity_no_gj %>%  
  filter(hb %in% "Scotland",
         admission_type %in% "Elective Inpatients") %>%
  ggplot(aes(x = date, y = episodes, col = specialty_name)) +
  geom_point() +
  geom_line()
```
Can 

This shows where 

```{r}
activity_no_gj %>%  
  filter(hb %in% "Scotland",
         specialty_name %in% "Intensive Care Medicine",
         admission_type != "All Inpatients and Day cases",
         admission_type != "Not Specified") %>% 
  ggplot(aes(x = date, y = episodes, col = admission_type)) +
  geom_point() +
  geom_line()
```

       

### Admission by specialty

Better can shiny this one and pick the specialty_name

#### KPI: Do people have longer lengths of stay from elective or emergency admissions?

### KPI: The whole system: from what’s coming through the front door in A&E and then how it filters through; how many emergency admissions are there, how many electives, how does this impact the lengths of stay people have in hospital?

take out "All Inpatients and Day cases" & "Not Specified" as duplicating and not really any there

##### Episodes by Admission
```{r}
activity_no_gj %>%  
  filter(hb %in% "Scotland",
         specialty_name %in% "Infectious Diseases",
         admission_type %in% c("Elective Inpatients", 
                               "Emergency Inpatients", 
                               "Transfers")) %>% 
  ggplot(aes(x = date, y = episodes, col = admission_type)) +
  geom_point() +
  geom_line()
```

##### Spells by Admission
```{r}
activity_no_gj %>%  
  filter(hb %in% "Scotland",
         specialty_name %in% "Infectious Diseases",
         admission_type %in% c("Elective Inpatients", 
                               "Emergency Inpatients", 
                               "Transfers")) %>% 
  ggplot(aes(x = date, y = spells, col = admission_type)) +
  geom_point() +
  geom_line()
```


```{r}
unique(activity_no_gj$admission_type)
```


```{r}
unique(activity_no_gj$specialty_name)
```



```{r}


activity_no_gj %>%  
  filter(hb %in% "Scotland",
         admission_type %in% "Emergency Inpatients",
         episodes > 3000) %>% 
 # group_by(episodes) %>% 
  ggplot(aes(x = date, y = episodes, col = specialty_name)) +
  geom_point() +
  geom_line()
```

```{r}
bigest_type <- activity_no_gj %>% 
  filter(hb %in% "Scotland",
         admission_type %in% "Emergency Inpatients") %>%
  group_by(specialty_name) %>% 
  summarise(count = sum(episodes)) %>% 
  arrange(desc(count)) %>% 
  head(3) %>% 
  ungroup()

bigest_type
```



```{r}

`%!in%` <- negate(`%in%`)

activity_no_gj %>% 
  filter(hb %in% "Scotland",
         admission_type %in% "Emergency Inpatients",
         specialty_name %!in% c(bigest_type$specialty_name)) %>%
  group_by(specialty_name) %>% 
  summarise(count = sum(episodes)) %>% 
  arrange(desc(count)) 


```


```{r}
avg_specialty <- activity_no_gj %>% 
  filter(hb %in% "Scotland",
         admission_type %in% "Emergency Inpatients",
         specialty_name %!in% c(bigest_type$specialty_name)) %>%
  group_by(date) %>% 
  summarise(avg = mean(episodes)) 

avg_specialty
```


```{r}
activity_no_gj %>% 
  filter(hb %in% "Scotland",
         admission_type %in% "Emergency Inpatients",
         specialty_name == bigest_type$specialty_name) %>%
  ggplot(aes(x = date, y = episodes)) +
  geom_point() +
  geom_line(aes(col = specialty_name)) +
  #geom_point(aes(x = avg_specialty$date, y = avg_specialty$avg, col = "red")) +
  geom_line(aes(x = avg_specialty$date, y = avg_specialty$avg), col = "red", size = 1.5)
```







```{r}
bigest_avg <- activity_no_gj %>% 
  filter(hb %in% "Scotland",
         admission_type %in% "Emergency Inpatients") %>%
  group_by(specialty_name) %>% 
  summarise(count = mean(episodes)) %>% 
  arrange(desc(count)) %>% 
  head(1)

bigest_avg
```

```{r}
activity_no_gj %>%  
  filter(hb %in% "Scotland",
         admission_type %in% "Elective Inpatients",
         filter()) %>% 
 # group_by(episodes) %>% 
  ggplot(aes(x = date, y = episodes, col = specialty_name)) +
  geom_point() +
  geom_line()
```


```{r}

```

